---
title: "lab week5: logistic regression"
author: "Robert van Timmeren s4535553"
date: "Generation date: `r format(Sys.time(), '%b %d, %Y - %H:%M:%S')`"
output: 
  html_document:
    toc: true
    code_folding: show
    toc_float: 
        collapsed: false
        smooth_scroll: true
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

packages <- c("rms", "mlogit")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())), repos='http://cran.us.r-project.org') 
}

library(rms) #needed for lrm()
library(mlogit) # needed for mlogit.data() and mlogit()
```


# Load data

We create a data frame from the table shown in this week's slides (slide 7, do not include column "both").

To avoid typing all the observations (there are 184!) we use the function "c()" to concatenate information and function "rep()" to repeat one element multiple times.

```{r, fig.width=5, fig.height=5}
# create the variables
status <- c(rep("upper", 36), rep("middle", 94), rep("lower", 54)) # exploratory variable / predictor
outcome <- c(rep("r", 30), rep("schwa", 6), rep("r", 20), rep("schwa", 74), 
             rep("r", 4), rep("schwa", 50)) # dependent variable / response

# create the data frame
ny.data <- data.frame(status, outcome)
table(ny.data) # this should show the same table as the one in slide 7
```


# Factor variables and reference levels

```{r, fig.width=5, fig.height=5}
ny.data$status <- factor(ny.data$status)
ny.data$outcome <- factor(ny.data$outcome)
str(ny.data)

ny.data$outcome <- relevel(ny.data$outcome, ref="schwa")
ny.data$status <- relevel(ny.data$status, ref="upper")
table(ny.data)
```


# Fit a binomial model

```{r, fig.width=5, fig.height=5}
glm_model <- glm(outcome ~ status, family = binomial, data = ny.data)
summary(glm_model)
```


# Assumptions

**Linearity:**

Predictor variable is not quantitative, so we do not check linearity. Therefore passes.

**Multicollinearity**:
```{r, fig.width=5, fig.height=5}
rms::vif(glm_model)
```

No values > 5, so passes multicollinearity assumption.


# Odds of "r" to "schwa" when the predictor status equals level "middle"

```{r, fig.width=5, fig.height=5}
glm_model$coefficients
exp(-1.3083) # intercept + status middle
```


# Probability of the outcome being "r" in this case, i.e. status "middle"?

```{r, fig.width=5, fig.height=5}
plogis(-1.3083)
```


# Load multinomial data

```{r, fig.width=5, fig.height=5}
status <- c(rep("upper", 68), rep("middle", 125), rep("lower", 71))

outcome <- c(rep("r", 30), rep("schwa", 6), rep("both", 32), rep("r", 20), rep("schwa", 74), 
             rep("both", 31), rep("r", 4), rep("schwa", 50), rep("both", 17)) # dependent

ny.data2 <- data.frame(status, outcome)
table(ny.data2)

ny.data2$status <- factor(ny.data2$status)
ny.data2$outcome <- factor(ny.data2$outcome)
str(ny.data2)
```


# Reference levels (multinomial)

```{r, fig.width=5, fig.height=5}
ny.data2$outcome <- relevel(ny.data2$outcome, ref="schwa")
ny.data2$status <- relevel(ny.data2$status, ref="upper")
levels(ny.data2$outcome)
levels(ny.data2$status)
```


# Fit a multinomial model

```{r, fig.width=5, fig.height=5}
library(mlogit)
multi.df <- mlogit.data(ny.data2, shape="wide", choice="outcome")
multi_model <- mlogit(outcome ~ 1 | status, data=multi.df, reflevel="schwa")
summary(multi_model)
```


# Odds of "both" to "schwa" when the status equals "upper"

```{r, fig.width=5, fig.height=5}
summary(multi_model)$coefficients
exp(1.673971)
```
