---
title: "lab 4 (week 5)"
author: "Robert van Timmeren (S4535553)"
date: "Generation date: `r format(Sys.time(), '%b %d, %Y - %H:%M:%S')`"
output: 
  html_document:
    toc: true
    code_folding: show
    toc_float: 
        collapsed: false
        smooth_scroll: true
    number_sections: true
---


```{r setup, message=F, echo=F}
knitr::opts_chunk$set(echo = TRUE)

packages <- c("car", "interactions", "rms")
if (length(setdiff(packages, rownames(installed.packages()))) > 0) {
  install.packages(setdiff(packages, rownames(installed.packages())), repos='http://cran.us.r-project.org') 
}

library(car)
library(interactions)
library(rms)
```


# Load the dataset

```{r load}
df = read.csv("lab4_cheese.csv")
head(df)
```


# Scatter plots and correlation coefficients

```{r, fig.width=5, fig.height=5}
# Acetic and H2S
plot(df$acetic ~ df$h2s, main = "Scatterplot Acetic and H2S")
abline(lm(df$acetic ~ df$h2s))
cor(df$acetic, df$h2s)

# Acetic and Lactic
plot(df$acetic ~ df$lactic, main = "Scatterplot Acetic and Lactic")
abline(lm(df$acetic ~ df$lactic))
cor(df$acetic, df$lactic)

# H2S and lactic
plot(df$h2s ~ df$lactic, main = "Scatterplot H2S and Lactic")
abline(lm(df$h2s ~ df$lactic))
cor(df$h2s, df$lactic)
```

# Multiple linear regression

```{r, fig.width=5, fig.height=5}
confint(lm(df$taste ~ df$acetic + df$h2s + df$lactic))
summary(lm(df$taste ~ df$acetic + df$h2s + df$lactic))$coefficients
```

Only H2S and lactic are significant. 

Acetic: p-value of 0.942 > 0.05 and therefore not significant. 
When acetic increases by 1, taste increases by 0.328.

H2S: p-value of 0.004 <= 0.05 and therefore significant.
When H2S increases by 1, taste increases by 3.912.

Lactic: p-value of 0.031 <= 0.05 and therefore significant.
When lactic increases by 1, taste increases by 19.671.


# Interactions

```{r, fig.width=5, fig.height=5}
# Acetic and H2S
m.int <- lm(taste ~ h2s * acetic + lactic, data=df)
summary(m.int)$coefficients

# H2S and lactic
m.int <- lm(taste ~ h2s * lactic + acetic, data=df)
summary(m.int)$coefficients

# Acetic and lactic
m.int <- lm(taste ~ acetic * lactic + h2s, data=df)
summary(m.int)$coefficients

# Lowest p-value is of acetic and lactic with p-value of 0.294.
library(jtools)
interact_plot(m.int, pred='lactic', modx='acetic')

```


# Assumption: linear relationship

```{r, fig.width=5, fig.height=5}

#par(mfrow = c(1, 3))
library(car)
m <- lm(taste ~ acetic + lactic + h2s, data = df)
crPlot(m, var = 'acetic')
crPlot(m, var = 'lactic')
crPlot(m, var = 'h2s')

# All of the variables seem fairly linear.
```


# Assumption: errors vary constantly (homoscedasticity)

```{r, fig.width=5, fig.height=5}
ncvTest(m)

# P-value > 0.05, so accept null hypothesis that the error has constant variance with the response.
```


# Assumption: multicollinearity and autocorrelation of residuals

```{r, fig.width=5, fig.height=5}
car::vif(m)

# No scores exceeds 5.

durbinWatsonTest(m)
acf(resid(m))

# P-value > 0.05, so accept null hypothesis that the value of the variable depends on the previous or next.
```


# Assumption: residuals normally distributed

```{r, fig.width=5, fig.height=5}
plot(m, which = 2)

# The residuals seem to be fairly normally distributed in the plot.

shapiro.test(residuals(m))$p.value

# P-value > 0.05, which means we accept the null hypothesis which states that it is normally distributed.


```


# Variable selection

```{r, fig.width=5, fig.height=5}
m0 <- lm(taste ~ 1, data = df)
m.fw <- step(m0, direction = 'forward', scope = ~ df$acetic + df$h2s + df$lactic)

# No, not all variables are retained. The best model consists of h2s and lactic.

m2 <- lm(df$taste ~ df$acetic + df$h2s + df$lactic)
m.bw <- step(m2, direction = 'backward', scope = ~df$acetic + df$h2s + df$lactic)

# Also conducting backward to see how this works (also possesses less risk). The outcome is the same, where the model consists of h2s and lactic.

# The model does fulfill the assumptions checked before, because all the variables in the model passed the assumptions.
```


# Overfitting

```{r, fig.width=5, fig.height=5}
m.val <- ols (taste ~ h2s + lactic, x = TRUE, y = TRUE, data = df)
val <- validate(m.val, bw=TRUE, B=1000, pr=FALSE)
val[1:5,]

# Model does not seem to be overfitted. There are no large optimism values (which indicate overfitting) and the slope optimism is far below 0.05. 
```
